---
title: "Geomorphic Morphometrics on canid skulls"
output: html_notebook
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE, fig.width = 8.5)
options(knitr.graphics.error = FALSE)
options(knitr.kable.NA = '')
options(rgl.useNULL=TRUE) # Note: OpenGL is depreciated on Macs and requires XQuartz in order to work properly. This will cause an error when loading the Morpho and geomorph packages. Setting this option removes the error. If you have XQuartz and don't mind it opening, suppress this line of code to regain full functionality.

library(tidyverse)
library(janitor)
library(gtools)
library(ggplot2)
library(patchwork)
library(kableExtra)
library(ggrepel)
library(geomorph)
# devtools::install_github('SlicerMorph/SlicerMorphR')
library(SlicerMorphR)
# devtools::install_github("lindsaywaldrop/munchcolors")
library(munchcolors)
```

## Running principal component and k-means clustering analyses

```{r}
dir.create("./results", showWarnings = F)
dir.create("./results/figures", showWarnings = F)
source("./src/GM_stats.R")
```

## Plotting PC and k-means results


```{r plotting-setup, include=FALSE}
# Setting UKC group order
UKC_group_order <- c("NATURAL", "Herding", "Terrier", "Companion", "Gun dog", 
                     "Guardian", "Northern Breed", "Sighthound", "Scenthound")
AKC_group_order <- c("NATURAL", "Herding", "Terrier", "Toy", "Sporting", 
                     "Working", "Non-sporting", "Hound")

# grouping plots ----------------------------------------------------------
# First added your ordered classifiers onto the PCA data
SlicerMorph.repc$UKC <- clsf_reord$UKC.breeding.Standard
SlicerMorph.repc$AKC <- clsf_reord$AKC.breeding.standard
# Set mixed breed to NA for plotting, there is only one!
SlicerMorph.repc$UKC[SlicerMorph.repc$UKC == "MIXED"] <- NA
SlicerMorph.repc$AKC[SlicerMorph.repc$AKC == "MIXED"] <- NA
# Setting fox to natural for AKC/UKC grouping to simplify plots
SlicerMorph.repc$UKC[SlicerMorph.repc$UKC == "FOX"] <- "NATURAL"
SlicerMorph.repc$AKC[SlicerMorph.repc$AKC == "FOX"] <- "NATURAL"

# Changing Yes to lowercase
SlicerMorph.repc$Nosework <- ifelse(clsf_reord$Nose =="Yes" | clsf_reord$Nose == "no", 
                                    tolower(clsf_reord$Nose), clsf_reord$Nose)
# Changing Nosework to ordered factor
SlicerMorph.repc$Nosework <- factor(SlicerMorph.repc$Nosework, ordered = T, 
                                    levels = c("yes", "no", "NATURAL", "FOX"))
# Changing Yes to lowercase 
SlicerMorph.repc$Bitework <- ifelse(clsf_reord$Bitework =="Yes" | clsf_reord$Bitework == "no",
                                    tolower(clsf_reord$Bitework), clsf_reord$Bitework)
# Changing Bitework to ordered factor
SlicerMorph.repc$Bitework <- factor(SlicerMorph.repc$Bitework, ordered = T, 
                                    levels = c("yes", "no", "NATURAL", "FOX"))
# Putting cluster levels into a column as a factor
SlicerMorph.repc$cluster <- factor(km$cluster)

# Now to ease viewing we'll factor them into broad categories

# Turning UKC isnto an ordered factor column
SlicerMorph.repc$UKC.shapefact <- factor(SlicerMorph.repc$UKC, ordered = T, 
                                         levels = UKC_group_order)

## AKC next
SlicerMorph.repc$AKC.shapefact <- factor(SlicerMorph.repc$AKC, 
                                         levels = AKC_group_order)

SlicerMorph.repc$AKC.fact <- factor(SlicerMorph.repc$AKC)

## Setting up a column that is domestic (dogs) and natural (foxes, others)
SlicerMorph.repc$domestic <- factor(ifelse(SlicerMorph.repc$UKC == "NATURAL" | 
                                             SlicerMorph.repc$UKC == "FOX", 
                                           "Natural", "Domesticated"), 
                                    ordered = T, levels = c("Domesticated", "Natural"))

## Calculating group means for setting skulls on plots: 
UKC <- SlicerMorph.repc[!is.na(SlicerMorph.repc$UKC),]
UKC <- UKC[,c("PC.1", "PC.2", "UKC")]
UKC_scenthound_mean <- data.frame(x = mean(UKC$PC.1[UKC$UKC == "Scenthound"]),
                           y = mean(UKC$PC.2[UKC$UKC == "Scenthound"]))
UKC_companion_mean <- data.frame(x = mean(UKC$PC.1[UKC$UKC == "Companion"]),
                                 y = mean(UKC$PC.2[UKC$UKC == "Companion"]))
AKC <- SlicerMorph.repc[!is.na(SlicerMorph.repc$AKC),]
AKC <- AKC[,c("PC.1", "PC.2", "AKC")]
AKC_hound_mean <- data.frame(x = mean(AKC$PC.1[AKC$AKC == "Hound"]),
                             y = mean(AKC$PC.2[AKC$AKC == "Hound"]))
AKC_nonsporting_mean <- data.frame(x = mean(AKC$PC.1[AKC$AKC == "Non-sporting"]),
                             y = mean(AKC$PC.2[AKC$AKC == "Non-sporting"]))
AKC_toy_mean <- data.frame(x = mean(AKC$PC.1[AKC$AKC == "Toy"]),
                             y = mean(AKC$PC.2[AKC$AKC == "Toy"]))
```

K-means clustering plot: 

```{r k-means, echo=FALSE, warning=FALSE, fig.cap="K-means clustering plot."}
#now build the plot
pkmean <- ggplot(SlicerMorph.repc[!is.na(SlicerMorph.repc$UKC),], aes(PC.1, PC.2, color = cluster, 
                                   fill = cluster,
                                   shape = domestic)) + 
  geom_point(size = 2) +
  stat_ellipse(geom = "polygon", aes(fill = cluster, group = cluster), level = 0.95, alpha = 0.1) +
  scale_shape_manual(values = c(5, 19), name = " ") +
  scale_color_munch(choice = "Nietzsche", discrete = TRUE,name = "Cluster") + 
  scale_fill_munch(choice = "Nietzsche", discrete = TRUE, name = "Cluster") +
  geom_text_repel(label = SlicerMorph.repc$Breed[!is.na(SlicerMorph.repc$UKC)], max.overlaps = 4, show.legend = FALSE)+
  xlab("PC 1 (Var = 50.2%)") + ylab("PC 2 (Var = 13.3%)") +
  theme_bw()
pkmean
```

Kennel-club groupings and PC results:

```{r kc-groups, echo=FALSE, warning=FALSE, fig.cap="K-means clustering plot.", fig.height=5.5, fig.width=12}
#now build the plot
kc_palette <- munch_palette("Murderer", 8)
kc_palette <- c(kc_palette, kc_palette[2])
pUKC <- ggplot(SlicerMorph.repc[!is.na(SlicerMorph.repc$UKC),], aes(PC.1, PC.2, color = UKC.shapefact, 
                                   fill = UKC.shapefact,
                                   shape = UKC.shapefact)) + 
  geom_point(size = 2) +
  stat_ellipse(geom = "polygon", aes(group = UKC.shapefact), level = 0.95, alpha = 0.2) +
  scale_shape_manual(values = c(19, 0, 1, 2, 5, 6, 7, 9, 10), name = "UKC Groups") +
  scale_color_manual(values = kc_palette, name = "UKC Groups") + 
  scale_fill_manual(values = kc_palette, name = "UKC Groups") +
  geom_text_repel(label = SlicerMorph.repc$Breed[!is.na(SlicerMorph.repc$UKC)], 
                  max.overlaps = 4, show.legend = FALSE) +
  annotate("segment", x = 0.1, xend = UKC_companion_mean$x, y = -0.1, yend = UKC_companion_mean$y,
             lty = 1, color = kc_palette[4]) +
  annotate("segment", x = -0.24, xend = UKC_scenthound_mean$x, y = -0.1, yend = UKC_scenthound_mean$y,
             lty = 1, color = kc_palette[9]) +
  xlim(-0.31, NA) +
  ylim(-0.20, NA) +
  xlab("PC 1 (Var = 50.2%)") + ylab("PC 2 (Var = 13.3%)") +
  theme_bw()
pAKC <- ggplot(SlicerMorph.repc[!is.na(SlicerMorph.repc$AKC),], aes(PC.1, PC.2, color = AKC.shapefact, 
                                   fill = AKC.shapefact,
                                   shape = AKC.shapefact)) + 
  geom_point(size = 2) +
  stat_ellipse(geom = "polygon", aes(fill = AKC.shapefact, 
                                     group = AKC.shapefact), level = 0.95, alpha = 0.2) +
  scale_shape_manual(values = c(19, 0, 1, 2, 5, 6, 7, 9, 10), name = "AKC Groups") +
  scale_color_munch(choice = "Murderer", discrete = TRUE, name = "AKC Groups") + 
  scale_fill_munch(choice = "Murderer", discrete = TRUE, name = "AKC Groups") +
  geom_text_repel(label = SlicerMorph.repc$Breed[!is.na(SlicerMorph.repc$AKC)], max.overlaps = 4, show.legend = FALSE)+
  annotate("segment", x = -0.22, xend = AKC_hound_mean$x, y = -0.05, yend = AKC_hound_mean$y,
             lty = 1, color = kc_palette[8]) +
  annotate("segment", x = 0.1, xend = AKC_nonsporting_mean$x, y = 0.15, yend = AKC_nonsporting_mean$y,
             lty = 1, color = kc_palette[7]) +
  annotate("segment", x = 0.16, xend = AKC_toy_mean$x, y = -0.07, yend = AKC_toy_mean$y,
             lty = 1, color = kc_palette[4]) +
  xlim(-0.31, NA) +
  ylim(NA, 0.20) +
  xlab("PC 1 (Var = 50.2%)") + ylab("PC 2 (Var = 13.3%)") +
  theme_bw() 

pUKC + pAKC
```


## Discriminant Factors between groups

Comparison of landmarks between group means

```{r}
# Extracting first 5 PCs and each kennel-club grouping scheme
skulls <- cbind(SlicerMorph.repc[, 1:5], 
                "UKC" = factor(SlicerMorph.repc$UKC, ordered = T, levels = UKC_group_order), 
                "AKC" = factor(SlicerMorph.repc$AKC, ordered = T, levels = AKC_group_order),
                "bitework" = factor(SlicerMorph.repc$Bitework),
                "scentwork" = factor(SlicerMorph.repc$Nosework))
skulls <- na.omit(skulls)

# Create a paired and ordered GPA-grouping object
mixdrop<-gpa$coords[,,order(dimnames(gpa$coords)[[3]])]
mixdrop<-mixdrop[,,-96]

reforge<-geomorph.data.frame(shape = mixdrop,
                             UKC = na.omit(factor(SlicerMorph.repc$UKC, ordered = T)),
                             AKC = na.omit(factor(SlicerMorph.repc$AKC, ordered = T)),
                             bitework = na.omit(factor(SlicerMorph.repc$Bitework[-96])),
                             scentwork = na.omit(factor(SlicerMorph.repc$Nosework[-96])))

# Create a function that subsets the mean landmarks for each group and does a pairwise subtraction and squaring of those means
meanforge <- function(target, group) {
  #sub-setting the means and sending them to a new object
  new.coords<-coords.subset(target[["shape"]], group = target[[group]])
  output_means<-lapply(new.coords, mshape)
  #perform the subtraction and squaring of the pairwise combinations
  result<- combn(output_means, 2, function(x) (x[[1]]-x[[2]])^2, simplify = FALSE)
  #rename the pairs so they're readable
  names(result)<-combn(names(output_means), 2, function(n) paste(n[1], "-", n[2]), simplify = TRUE)
  return(result)
}

#now run the function for each grouping scheme. Each member of the list will be named based on which two groups are compared
meanshape.UKC<-meanforge(reforge,"UKC")
meanshape.AKC<-meanforge(reforge,"AKC")
meanshape.nose<-meanforge(reforge,"scentwork")
meanshape.bite<-meanforge(reforge,"bitework")

```

Reformatting data for visualization:

```{r}
calc_gpa_dist <- function(dat){
  require(tidyr)
  n <- length(names(dat))
  results <- matrix(NA, nrow = nrow(dat[[1]]), ncol = n)
  results <- as.data.frame(results)
  results <- cbind(seq(1,nrow(results)), results)
  colnames(results) <- c("landmark",names(dat))
  
  for(i in names(dat)){
    results[[i]] <- sqrt(dat[[i]][,"X"]^2 + dat[[i]][,"Y"]^2 + dat[[i]][,"Z"]^2)
  }
  results_long <- pivot_longer(results, cols = -"landmark", names_to = "comparison")
  results_long$landmark <- factor(results_long$landmark)
  return(results_long)
}

meanshape.AKC[["plotting"]] <- calc_gpa_dist(meanshape.AKC)
meanshape.UKC[["plotting"]] <- calc_gpa_dist(meanshape.UKC)
meanshape.bite[["plotting"]] <- calc_gpa_dist(meanshape.bite)
meanshape.bite[["cutoff"]] <- data.frame("mean" = mean(meanshape.bite[["plotting"]]$value), 
                                        "sd" = sd(meanshape.bite[["plotting"]]$value))
meanshape.nose[["plotting"]] <- calc_gpa_dist(meanshape.nose)
meanshape.nose[["cutoff"]] <- data.frame("mean" = mean(meanshape.nose[["plotting"]]$value), 
                                        "sd" = sd(meanshape.nose[["plotting"]]$value))

meanshape.AKC[["plotting"]]$short <- ifelse(grepl("toy", meanshape.AKC[["plotting"]]$comparison, 
                                                  ignore.case = T),
                                           "toy", 
                                           ifelse(grepl("natural", meanshape.AKC[["plotting"]]$comparison,
                                                        ignore.case = T),
                                                  "natural", "other"))
meanshape.AKC[["plotting"]]$short <- factor(meanshape.AKC[["plotting"]]$short, 
                                            ordered = T, levels = c("toy", "natural", "other"))
meanshape.AKC[["cutoff"]] <- data.frame("mean" = mean(meanshape.AKC[["plotting"]]$value), 
                                        "sd" = sd(meanshape.AKC[["plotting"]]$value))
meanshape.UKC[["plotting"]]$short <- ifelse(grepl("companion", meanshape.UKC[["plotting"]]$comparison,
                                                      ignore.case = T),
                                           "companion", 
                                           ifelse(grepl("natural", meanshape.UKC[["plotting"]]$comparison,
                                                      ignore.case = T),
                                           "natural", "other"))
meanshape.UKC[["plotting"]]$short <- factor(meanshape.UKC[["plotting"]]$short, 
                                            ordered = T, levels = c("companion", "natural", "other"))
meanshape.UKC[["cutoff"]] <- data.frame("mean" = mean(meanshape.UKC[["plotting"]]$value), 
                                        "sd" = sd(meanshape.UKC[["plotting"]]$value))

```

AKC & UKC groups:
```{r}
p_UKC_compare <- ggplot(meanshape.UKC[["plotting"]], aes(landmark, value, color = short)) + 
  geom_point(position = position_jitter(width = 0.1)) + 
  geom_hline(yintercept = meanshape.UKC[["cutoff"]]$mean+2*meanshape.UKC[["cutoff"]]$sd,
             lty = 2) +
  scale_color_viridis_d(name = "Comparison\nagainst:") +
  annotate("segment", x = 1, xend = 20, y = 0.0085, yend = 0.0085,
             lty = 1, color = "red") +
  annotate("text", x = 10, y = 0.0075, label = "Dorsal features", color = "red") +
  annotate("segment", x = 21, xend = 38, y = 0.0085, yend = 0.0085,
             lty = 1, color = "blue") +
  annotate("text", x = 30, y = 0.0075, label = "Ventral features", color = "blue") +
  ylab("Group-mean landmark\ndistance squared") +
  ggtitle("UKC group comparisons") +
  theme_bw() #+
  #theme(legend.position = "bottom")
  #theme(legend.position = "inside", legend.position.inside = c(0.75,0.65),
  #      legend.background = element_rect(color = "gray30", fill = "white", linetype="solid",
  #                                       linewidth = 0.2))

p_AKC_compare <- ggplot(meanshape.AKC[["plotting"]], aes(landmark, value, color = short)) + 
  geom_point(position = position_jitter(width = 0.1)) + 
  geom_hline(yintercept = meanshape.AKC[["cutoff"]]$mean+2*meanshape.AKC[["cutoff"]]$sd,
             lty = 2) +
  scale_color_viridis_d(name = "Comparison\nagainst:") +
  annotate("segment", x = 1, xend = 20, y = 0.0085, yend = 0.0085,
             lty = 1, color = "red") +
  annotate("text", x = 10, y = 0.0075, label = "Dorsal features", color = "red") +
  annotate("segment", x = 21, xend = 38, y = 0.0085, yend = 0.0085,
             lty = 1, color = "blue") +
  annotate("text", x = 30, y = 0.0075, label = "Ventral features", color = "blue") +
  ylab("Group-mean landmark\ndistance squared") +
  ggtitle("AKC group comparisons") +
  theme_bw() #+
  #theme(legend.position = "bottom")
    #theme(legend.position = "inside", legend.position.inside = c(0.75,0.65),
    #    legend.background = element_rect(color = "gray30", fill = "white", linetype="solid",
    #                                     linewidth = 0.2))
p_AKC_compare / p_UKC_compare
```
Nosework and Bitework: 

```{r}
p_bite_compare <-  ggplot(meanshape.bite[["plotting"]], aes(landmark, value, color = comparison)) + 
  geom_point(position = position_jitter(width = 0.1)) + 
  geom_hline(yintercept = meanshape.bite[["cutoff"]]$mean+2*meanshape.bite[["cutoff"]]$sd,
             lty = 2) +
  annotate("segment", x = 1, xend = 20, y = 0.007, yend = 0.007,
             lty = 1, color = "red") +
  annotate("text", x = 10, y = 0.0065, label = "Dorsal features", color = "red") +
    annotate("segment", x = 21, xend = 38, y = 0.007, yend = 0.007,
             lty = 1, color = "blue") +
  annotate("text", x = 30, y = 0.0065, label = "Ventral features", color = "blue") +
  scale_color_viridis_d(name = "Group-group\ncomparison") +
  ylab("Group-mean landmark\ndistance squared") +
  ggtitle("Bite-work group comparisons") +
  theme_bw()
p_nose_compare <- ggplot(meanshape.nose[["plotting"]], aes(landmark, value, color = comparison)) + 
  geom_point(position = position_jitter(width = 0.1)) + 
  geom_hline(yintercept = meanshape.nose[["cutoff"]]$mean+2*meanshape.nose[["cutoff"]]$sd,
             lty = 2) +
  annotate("segment", x = 1, xend = 20, y = 0.007, yend = 0.007,
             lty = 1, color = "red") +
  annotate("text", x = 10, y = 0.0065, label = "Dorsal features", color = "red") +
    annotate("segment", x = 21, xend = 38, y = 0.007, yend = 0.007,
             lty = 1, color = "blue") +
  annotate("text", x = 30, y = 0.0065, label = "Ventral features", color = "blue") +
  scale_color_viridis_d(name = "Group-group\ncomparison") +
  ylab("Group-mean landmark\ndistance squared") +
  ggtitle("Scent-work group comparisons") +
  theme_bw()

p_bite_compare / p_nose_compare + plot_layout(guides = "collect")
```

## MANOVA 

```{r}
man_UKC <- manova(as.matrix(skulls[,1:5])~skulls$UKC)
summary.aov(man_UKC, test="Pillai")

#using Geomorph and doing pairwise comparisons
manUKC<-procD.lm(shape ~ UKC, data = reforge, RRPP = TRUE)
manAKC<-procD.lm(shape ~ AKC, data = reforge, RRPP = TRUE)
manBite<-procD.lm(shape ~ bitework, data = reforge, RRPP = TRUE)
manScent<-procD.lm(shape ~ scentwork, data = reforge, RRPP = TRUE)

hold.ukc<-manova(manUKC)
hold.akc<-manova(manAKC)
hold.bite<-manova(manBite)
hold.nose<-manova(manScent)

ukcpairs<-pairwise(manUKC, groups = reforge$UKC)
akcpairs<-pairwise(manAKC, groups = reforge$AKC)
bitepairs<-pairwise(manBite, groups = reforge$bitework)
nosepairs<-pairwise(manScent, groups = reforge$scentwork)


pvals.UKC<-summary(ukcpairs, test.type= "dist", confidence = 0.95, stat.table = TRUE)
pvals.UKC<-pvals.UKC$pairwise.tables$P[lower.tri(pvals.UKC$pairwise.tables$P)]
pvals.AKC<-summary(akcpairs, test.type= "dist", confidence = 0.95, stat.table = TRUE)
pvals.AKC<-pvals.AKC$pairwise.tables$P[lower.tri(pvals.AKC$pairwise.tables$P)]
pvals.bite<-summary(bitepairs, test.type= "dist", confidence = 0.95, stat.table = TRUE)
pvals.bite<-pvals.bite$pairwise.tables$P[lower.tri(pvals.bite$pairwise.tables$P)]
pvals.nose<-summary(nosepairs, test.type= "dist", confidence = 0.95, stat.table = TRUE)
pvals.nose<-pvals.nose$pairwise.tables$P[lower.tri(pvals.nose$pairwise.tables$P)]

pvals.ukc.adjusted<-p.adjust(pvals.UKC, method = "bonf")
pvals.akc.adjusted<-p.adjust(pvals.AKC, method = "bonf")
pvals.bite.adjusted<-p.adjust(pvals.bite, method = "bonf")
pvals.nose.adjusted<-p.adjust(pvals.nose, method = "bonf")
```

```{r}
lda_UKC <- MASS::lda(skulls$UKC~as.matrix(skulls[,1:5]))
lda_UKC
```

## Testing of Functional Groups

```{r fxn-groupings, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3.5, fig.cap="Bitework and scent work."}
fxn_group_palette <- munch_palette("Murderer",8)[c(1:3,5)]
pbite <- ggplot(SlicerMorph.repc, aes(PC.1, PC.2, color = Bitework, 
                                   fill = Bitework,
                                   shape = Bitework)) + 
  geom_point() +
  stat_ellipse(geom = "polygon", level = 0.95, alpha = 0.1) +
  scale_shape_manual(values = c(21, 23, 0, 2), name = " ") +
  scale_color_manual(values = fxn_group_palette, name = " ") + 
  scale_fill_manual(values = fxn_group_palette, name = " ") +
  xlab("PC 1 (Var = 50.2%)") + 
  ylab(" ") +
  ggtitle("Selected for bite work") +
  theme_bw() + theme(legend.position = "right")
pscent <- ggplot(SlicerMorph.repc, aes(PC.1, PC.2, color = Nosework, 
                                   fill = Nosework,
                                   shape = Nosework)) + 
  geom_point() +
  stat_ellipse(geom = "polygon", level = 0.95, alpha = 0.1) +
  scale_shape_manual(values = c(21, 23, 0, 2), name = " ") +
  scale_color_manual(values = fxn_group_palette, name = " ") + 
  scale_fill_manual(values = fxn_group_palette, name = " ") +
  xlab("PC 1 (Var = 50.2%)") + 
  ylab("PC 2 (Var = 13.3%)") +
  ggtitle("Selected for scent work") +
  theme_bw() 
pscent + pbite + plot_layout(guides = "collect") & theme(legend.position = "right")

```

```{r bite-force, echo=FALSE, message=FALSE, warning=FALSE}
bite_data <- data.frame("Breed" = SlicerMorph.repc$Breed, 
                        "Bitework" = SlicerMorph.repc$Bitework, 
                        "Domestic" = SlicerMorph.repc$domestic, 
                        "UKC" = SlicerMorph.repc$UKC, 
                        "AKC" = SlicerMorph.repc$AKC,
                        "canine" = SlicerMorph.repc$bfq_canine, 
                        "carnassial" = SlicerMorph.repc$bfq_carnassial)
bite_data_long <- pivot_longer(bite_data, cols = c("canine","carnassial"))
pforce <- ggplot(bite_data_long, aes(Bitework, value, fill = name)) + 
  geom_boxplot(alpha = 0.6, outlier.shape = NA) + 
  geom_point(alpha = 0.5, position = position_jitterdodge(jitter.width = 0.2), shape = 21) +
  scale_fill_manual(values = munch_palette("YellowLog",2), name = " ") +
  ylab("Bite-force quotient") + xlab("Selected for bite work") +
  theme_bw()
pforce
```

```{r bite-force-stats}
bite_yes_canine <- bite_data$canine[bite_data$Bitework == "yes"]
bite_no_canine <- bite_data$canine[bite_data$Bitework == "no"]
shapiro.test(bite_yes_canine)
shapiro.test(bite_no_canine)
t.test(bite_yes_canine, bite_no_canine)

bite_yes_carnassial <- bite_data$carnassial[bite_data$Bitework == "yes"]
bite_no_carnassial <- bite_data$carnassial[bite_data$Bitework == "no"]
shapiro.test(bite_yes_carnassial)
shapiro.test(bite_no_carnassial)
t.test(bite_yes_carnassial, bite_no_carnassial)

bite_natural_canine <- bite_data$canine[bite_data$Bitework == "NATURAL"]
shapiro.test(bite_natural_canine)
bite_fox_canine <- bite_data$canine[bite_data$Bitework == "FOX"]
shapiro.test(bite_fox_canine)

t.test(bite_yes_canine, bite_natural_canine)
t.test(bite_yes_canine, bite_fox_canine)
```

## For the paper: 

Figure 2: k-means cluster on PC plot

```{r fig-2, echo=FALSE, warning=FALSE, fig.cap="A. K-means clustering plot.", fig.height=4, fig.width=9.6}
pkmean + xlim(-0.22, 0.25) + theme(legend.margin=margin(c(0,0,0,0)))
ggsave("./results/figures/figure2.pdf", height = 4, width = 9.6)
```

Figure 3: UKC comparison figure

```{r fig-3, echo=FALSE, warning=FALSE, fig.cap="A. K-means clustering plot. B. AKC groupings. C. UKC groupings.", fig.height=7.5, fig.width=12}
design2 <- 
  "AAAA
   AAAA
   BBBB"
(pUKC ) / 
  free(p_UKC_compare + ggtitle(" ") +  
         theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))) + 
  plot_annotation(tag_levels = "A") +
  plot_layout(design = design2) & theme(legend.margin=margin(c(0,0,0,0)))
  
ggsave("./results/figures/figure3.pdf", height = 3.45*2, width = 3.25*2)
```

Figure 4: AKC comparison figure

```{r fig-4, echo=FALSE, warning=FALSE, fig.cap=" AKC groupings. ", fig.height=7.5, fig.width=12}
design2 <- 
  "AAAA
   AAAA
   BBBB"
pAKC / free(p_AKC_compare + ggtitle(" ") + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)))+
  plot_annotation(tag_levels = "A") +
  plot_layout(design = design2) & theme(legend.margin=margin(c(0,0,0,0)))
ggsave("./results/figures/figure4.pdf", height = 3.45*2, width = 3.25*2)
```

Figure 5: task-specific group results

```{r fig-5, echo=FALSE, warning=FALSE, fig.cap="A. K-means clustering plot. B. AKC groupings. C. UKC groupings.", fig.height=7.5, fig.width=12}
design2 <- 
  "AAAA
   BBBB
   BBBB
   CCCC"
(pscent + pbite + plot_layout(guides = "collect") & theme(legend.position = "right")) / 
  (p_nose_compare /p_bite_compare + 
     plot_layout(guides = "collect") & theme(legend.position = "right", 
                                             axis.text.x = element_text(angle = 90, 
                                                                        vjust = 0.5, hjust=1))) / 
  pforce  + 
  plot_annotation(tag_levels = "A") +
  plot_layout(design = design2) & theme(legend.margin=margin(c(0,0,0,0)))
ggsave("./results/figures/figure5.pdf", width = 9.6, height = 11.52)
```






